---
title: "Untitled"
author: "Wang Shufan"
date: "2024-03-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(tidyverse,targets,readxl,gt)
```

### Effect of treatment on the effect of growth factor on the gene expression
This project aims to analysis how the difference treatments influence the growth factor? and the gene expression.

The data is provided by Karl Beartor.

We will output the statis and plots that can show relationship, and display them in some slides.

### 2024-03-04

## Problem 1: the aim of the project
Analyze the effect of the treatment on the gene expression or on the growth factor? Or analyze the effect of the growth factor of the gene expression?

NO groth factor in the data = gene expression?

### Data
The csv dataset includes 88 obversations and 5 variables. 

cell-line: categorical variables with values "Wild-type", "Cell-type 101", "WILD-TYPE" and "CELL-TYPE 101".

treatment: categorical variables with values "Placebo" and "Activating factor 42".

name: categorical variables with 8 possible values. do not know what it represents for.

conc: concentration. numerical variables. with integer values ranges from 0 to 10.

gene_expression: numerical variables. response variables


no missing values

## Question 2: The values in cell-line
Are "wild-type" and "WILD-TYPE" the same type? 
Are "Cell-type 101" and "CELL-TYPE" the same type?

## Question 3: The meaning of the name variable
What does the "name" variable represent for?

## Question 4: The concentration variable.
Do the values of the "conc" variable have 

### 2024-3-3 EDA for raw data

The EDA of the raw data has been done to see if the data needed to be cleaned. All the operations's results were stored in this list.

```{r}
raw_EDA_list <- tar_read(gene_raw_EDA)
```



I firstly checked the summary of the raw data.

The raw data have 88 observations and 5 variables: 3 categoritical variables (cell_line, treatment, name) and 2 numerical variables(conc, gene_expression)

```{r}
raw_EDA_list$summary
```
According to @fig-missing-value, there is no missing values in the dataset.
```{r}
#| fig-cap: Missing values plot
#| label: fig-missing-value
raw_EDA_list$missing_plot
```


There are two class in the cell_line: "Cell-type 101" and "Wild-type".
We need to fix the typos.
```{r}
raw_EDA_list$cell_line_count
```

2 classes in treatment: Activating factor 42" and "Placebo".
Need to fix the typos.
```{r}
raw_EDA_list$treatment_count
```

8 classes in name.
Need to fix the typos.
```{r}
raw_EDA_list$name_count
```
According to @fig-histgrow, all 8 classes of cells have been tested with 11 different growth factors. No missing tests. :D
```{r}
#| fig-cap: Histogram of growth factor.
#| label: fig-histgrow
raw_EDA_list$conc_histgram
```

According to @fig-histgenexp, the distribution of gene expression is unimodal and right-skewed. The majority of observations had gene expression <25. After checking with Karl, the gene expression > 25 are not outliers
```{r}
#| fig-cap: Histogram of gene expression.
#| label: fig-histgenexp
raw_EDA_list$expression_histgram
```

## 2024-3-10 Data clean

We cleaned the data by fixing all the typos in the dataset.
```{r}
tar_load(gene_clean)
head(gene_clean)
```
## 2024-3-11 EDA for cleaned data

In @fig-coplot, every subplot represents how gene expression changes with the values of growth factors for different names.  
The name strongly influenced the relationship between gene expression and growth factors.


```{r}
#| fig-cap: Conditioning plots of gene expression vs growth factor with different name
#| label: fig-coplot
colnames(gene_clean)[4]<- c('growth_factor')
coplot(gene_expression ~ growth_factor|name,data=gene_clean)
```

```{r}
tar_load(gene_clean_EDA)
```


Scatter plots of gene expression and growth factor are drawn.

Both @fig-scatter_wild and @fig-scatter_cell indicated that the strong linear relationships between gene expression and growth factor were greatly influenced by the names, which is consistenty with @fig-coplot. Moreover, the effects of treatment "Activating factor 42" on gene expression were also varied with different name. A mixed-effect model with random intercept and coefficients may fit the data well.

```{r}
#| fig-cap: Scatter plot of gene expression vs growth factor with different treatments,names and cell line Wild-type
#| label: fig-scatter_wild
gene_clean_EDA$scatter_Gene_vs_Conc_wild
```
```{r}
#| fig-cap: Scatter plot of gene expression vs growth factor with different treatments,names and cell line Cell-type 101
#| label: fig-scatter_cell
gene_clean_EDA$scatter_Gene_vs_Conc_cell
```




@fig-boxplot illustrated the median values and IQR of the gene expression with different growth factor and treatments. The introduction of boxplot can be found in 
Variations of box plots [@McGill1978, pp.12-16].

For median values:

1) Median increased as the growth factors increased.

2) “Activating factor 42” significantly outperformed the “Placebo” for all growth factors.

For IQR:

1) IQR increased as the growth factors increased.

2) For different treatment, the IQRs of the same growth factor are similar.
```{r}
#| fig-cap: Boxplots of gene expression vs growth factor with different treatments
#| label: fig-boxplot
gene_clean_EDA$boxplot_Gene_vs_Conc
```

@tab-mean displayed the mean of gene expression with different growth factors.

1) Mean grew as the growth factors increased in general.
2) "Activating factor 42" led to a larger mean than "Placebo".
```{r}
#| tbl-cap: Table of mean gene expression under various pairs of growth factor and treatement
#| label: tab-mean
gene_clean_EDA$mean_GeneExp
```
@tab-sd displayed the standard deviation of gene expression with different growth factors.

1) Variance grew as the growth factor increased in general.
2) "Activating factor 42" led to a larger variance than "Placebo".


```{r}
#| tbl-cap: Table of gene expression stadard deviations under various pairs of growth factor and treatement
#| label: tab-sd
gene_clean_EDA$sd_GeneExp
```

# 2024-4-3 Conference plot

We modified @fig-scatter_wild and @fig-scatter_cell, making them fit the requirements of the conference. The plot submitted for conference is a tiff file (9in x 6in) with a resolution of 500. The font in the figures should be Time New Roman. You can find the file submitted for the conference in 
```{r,echo=FALSE}
here::here("figs")
```


In this README file, for easier displaying, we used the default font type and size in 
```{r}
citation("ggplot2")
```

```{r}
tar_read(conference_plot)
```


# 2024-5-22 Sample Size
```{r}
tar_load(sample_size)
sample_size
```
# 2024-5-23 Models

We built 5 mix-effect linear model for the gene data. The response variable is gene_expression and each model had different the predictors and random effect. We selected the model with smallest AIC and significantly better than the other simper models, which is determined by p-values of Likelihood ratio tests.

```{r}
tar_load(gene_models)
```

Model 1
predictors:  conc and cell_line;
random effect: random intercepts for name;
```{r}
gene_models$me1
```
Model 2:
predictors: treatment, conc and cell_line;
interaction term: treat and conc;
random effect: random intercepts for name;
```{r}
gene_models$me2
```
Model 3:
predictors: treatment, conc and cell_line;
interaction term: treatment and conc
random effect: random intercepts for name;
```{r}
gene_models$me3
```
Model 4:
predictors: treatment, conc;
interaction term: treatment and conc;
random effect: name had random intercepts and slopes for conc and treatment were influenced by name.
```{r}
gene_models$me4
```
Model 4:
predictors: treatment, conc and cell line;
interaction term: treatment and conc;
random effect: name had random intercepts and slopes for conc and treatment were influenced by name.
```{r}
gene_models$me5
```

# 2024-5-29 Model Selection

According to @tab-anova, "me4" had the smallest AIC. Except for "me5", all other models are significantly better than their previous models. Therefore, we selected "me4" as out final model.
```{r}
#| tbl-cap: Table of AIC and p-values of Likelihhod ratio test
#| label: tab-anova
gene_models$anova
```
# 2024-5-29 Model Visulization
@fig-finalm illustrated the final model. According to the table, the observations with the same name evenly distributed around their related sublinear model, indicating our final model's goodness of fit. 
```{r}
#| fig-cap: Final model
#| label: fig-finalm
gene_models$final_model
```

@tab-coef_final displayed the intercepts and coefficients of sublinear models plotted in @fig-finalm.
```{r}
#| tbl-cap: The table of intercept and coefficients of the final model.
#| label: tab-coef_final
gene_models$coe_final
```
README file path:
```{r}
here::here("README.qmd")
```
Project path
```{r}
here::here()
```




